---
title: "Assignment3A"
author: "Cecilia De La Fuente Chavez"
format: html
execute:
  echo: true
editor: visual
---

## Approach Deliverable

### Overview

For Assignment 3A, to implement a Global Baseline Estimator, I will first look at the formula provided in the Excel sheet. The formula is Global Baseline Estimate = Mean Movie Rating + A movie's rating relative to average + A respondent's rating relative to average. The plan is to first establish the average rating across all movies and survey respondents, which if I remember correctly, I had already calculated in Assignment 2A, then each movie's rating relative to average and each respondent's rating relative to average. From here I can estimate the rating a particular user would give a movie, and if they have rated them I can compare their actual rating with the estimate.

## Code Base Deliverable

### Load libraries

```{r}
library(DBI)
library(RPostgres)
library(tidyverse)
```

### Connect to PostgreSQL

```{r}
con <- dbConnect(
  RPostgres::Postgres(),
  dbname = "607",
  host = "localhost",
  port = 5432,
  user = "postgres",
  password = Sys.getenv("POSTGRES_PASSWORD")
)
```

Checked current tables in database, assigned names in R

```{r}
dbListTables(con)
rating_lookup <- dbReadTable(con, "rating_lookup")
survey_responses <- dbReadTable(con, "survey_responses")
```

Transformed data frame so that each movie name became a value under movie column with their respective response value

```{r}
movie_ratings <- survey_responses %>%
  pivot_longer(
    cols = -id,
    names_to = "movie",
    values_to = "response"
  )
glimpse(movie_ratings)
```

Joined the rating_lookup table by response to add the corresponding numeric value to the ratings

```{r}
movie_ratings <- movie_ratings %>%
  left_join(rating_lookup, by = "response")
glimpse(movie_ratings)
```

Calculate the global mean excluding NAs

```{r}
global_mean <- mean(movie_ratings$rating, na.rm = TRUE)
global_mean
```

Create a new data frame that groups by movie and calculates the average rating for each on a new column movie_mean

```{r}
movie_effect <- movie_ratings %>%
  group_by(movie) %>%
  summarise(
    movie_mean = mean(rating, na.rm = TRUE)
  )
```

Calculate the relative rating to the global mean for each

```{r}
movie_effect <- movie_effect %>%
  mutate(
    movie_relative = movie_mean - global_mean
  )
```

Create a new data frame that groups by id and calculates the average rating each id gave on a new column user_mean

```{r}
user_effect <- movie_ratings %>%
  group_by(id) %>%
  summarise(
    user_mean = mean(rating, na.rm = TRUE)
  )
```

Calculate the relative rating to the global mean for each, but deal with NAs by assigning them a 0 rating

```{r}
user_effect <- user_effect %>%
  mutate(
    user_relative = user_mean - global_mean,
    user_relative = ifelse(is.nan(user_relative), 0, user_relative)
  )
```

Join both movie and user data frames to consolidate rating and all relative calculations

```{r}
baseline_data <- movie_ratings %>%
  left_join(
    movie_effect %>% select(movie, movie_relative),
    by = "movie"
  ) %>%
  left_join(
    user_effect %>% select(id, user_relative),
    by = "id"
  )
```

Calculate an rating estimate for each id and movie but cap rating at 5

```{r}
baseline_data <- baseline_data %>%
  mutate(
    baseline_estimate = global_mean + movie_relative + user_relative,
    baseline_estimate = pmin(baseline_estimate, 5)
  )
```

Make a subset of data frame called recommendations that filters for "Have not seen it" or NA responses and shows the estimate rating.

```{r}
recommendations <- baseline_data %>%
  filter(is.na(rating)) %>%
  arrange(id, desc(baseline_estimate))
```

### Conclusion

I used the survey ratings to calculate the global average rating, each movie's rating and each user's rating compared to the global average. I then combined these values to estimate how each user might rate movies they had not seen.
